---
title: "Data 606 Lab 3"
author: "John Perez"
date: "2/25/2019"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    css: ./lab.css
---


```{r load-data, eval=TRUE, include=FALSE}
load("more/bdims.RData")
mdims <- subset(bdims, sex == 1)
fdims <- subset(bdims, sex == 0)
fhgtmean <- mean(fdims$hgt)
fhgtsd   <- sd(fdims$hgt)
```

1.  Make a histogram of men's heights and a histogram of women's heights.  How 
    would you compare the various aspects of the two distributions?  

**Both distributions look fairly normal with slight skew.**
```{r}
par(mfrow=c(1,2))
hist(mdims$hgt, main="Men's Height")
hist(fdims$hgt,main="Women's Height")
```


2.  Based on the this plot, does it appear that the data follow a nearly normal 
    distribution?  

**Yes, the data appears to follow a nearly normal distribution** 

3.  Make a normal probability plot of `sim_norm`.  Do all of the points fall on 
    the line?  How does this plot compare to the probability plot for the real 
    data?

**This plot looks similar to the plot for the real data, having points that closely follow the line, but with points deviating from the line towards the tails.**  
    
```{r}
set.seed(2112)
sim_norm <- rnorm(n = length(fdims$hgt), mean = fhgtmean, sd = fhgtsd)
qqnorm(sim_norm)
qqline(sim_norm)
```

4.  Does the normal probability plot for `fdims$hgt` look similar to the plots 
    created for the simulated data?  That is, do plots provide evidence that the
    female heights are nearly normal?  

**Yes, the plots look similar to the plots created by the simulated data**

5.  Using the same technique, determine whether or not female weights appear to 
    come from a normal distribution.

**It appears weight is skewed right, diverging from the normal distribution**

```{r}
set.seed(22)
qqnormsim(fdims$wgt)
```

6.  Write out two probability questions that you would like to answer; one 
    regarding female heights and one regarding female weights.  Calculate the 
    those probabilities using both the theoretical normal distribution as well 
    as the empirical distribution (four probabilities in all).  Which variable,
    height or weight, had a closer agreement between the two methods?  
    
**What is the probability that a random female is shorter than 150cm?**  
**What is the probability that a random female is heavier than 68kg?**
```{r, eval=TRUE}
pnorm(150, fhgtmean, fhgtsd)
fwgtmean <- mean(fdims$wgt)
fwgtsd   <- sd(fdims$wgt)
1 - pnorm(68, fwgtmean, fwgtsd)
```
* * *

## On Your Own

-   
**a.** This corresponds to plot B  
**b.** This corresponds to plot C  
**c.** This corresponds to plot D  
**d.** This corresponds to plot A  

-   Note that normal probability plots C and D have a slight stepwise pattern.  
    Why do you think this is the case?
    **This could be caused by values being rounded**

-   Make a normal probability plot for female 
    knee diameter (`kne.di`).  Based on this normal probability plot, is this 
    variable left skewed, symmetric, or right skewed?  Use a histogram to confirm 
    your findings.
    **This variable is right skewed**
```{r}
#fkn <- fdims$kne.di
qqnorm(fdims$kne.di)
qqline(fdims$kne.di)
hist(fdims$kne.di)
```
    


